Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics

### 2020-01-14 paper/paper.Rmd
rearrange

The paper is organised as follows. The next section discusses the background of geographic data display and visual inference procedures. The [Methodology] Section describes the methods for conducting the experiment and analysing the results. The results are summarized in Section [Results]. 
Background
=============

@@

<!-- Specific problem for Australia 
Australia is an extreme case of heterogeneous geographic units, with the large differences between the smallest and largest geographic units across various granularities.
To communicate information regarding the small inner-city geographic units of Australian cities, an alternative display must be employed. This will emphasize the value of the statistic for areas with a high population. Fig. \ref{fig:liver} shows the distribution of the smoothed average amount of liver cancer diagnoses for Australian Males from 2005 to 2014.
These inner-city areas are not visible on the choropleth map (a). The hexagon tile map (b) shows orange hexagons for the inner-city SA2s with higher than average levels of diagnoses in the capital cities of Melbourne and Sydney. The northern Queensland and Northern Territory SA2s with higher than average rates are still visible but use less map space on the hexagon tile map. There are still many light blue areas in both displays. 
-->

@@

<!--We suggest the hexagon tile map as an alternative display. To test the effectiveness of these displays, there is a formal testing framework. Visual inference considers the communication of data or data summaries through visualizations. It considers these visualizations to be visual statistics, with observable features.
-->

@@

<!-- The lineup protocol was formalized as a test for the effectiveness of a visualization. Rather than classical statistical tests, it tasks human participants with finding the display that shows structure between the variables. If they do not detect the structure, then it is not significantly different from the null data.

Visual inference uses the general hypothesis structure:

 - Null hypothesis: There is \bf{no} relationship between the variables
 - Alternative hypothesis: There is a relationship between the variables

To test these hypotheses, the line up protocol randomly places a "guilty" data visualization in a lineup of "innocents". Where the guilty data set contains structure, and the innocents data sets do not.
In a grid of visualizations, an observer is asked to pick the display that is most different, if they select the data set containing structure, they have identified the guilty hidden within the  innocents [@GTPCCD].
The participants identify the guilty data as different from the innocent data with probability $1/m$, where $m$ is the number of null plots plus 1 to account for the guilty data set. When the guilty data set is chosen we reject the null hypothesis that it was innocent is rejected with a $1/m$ chance or type I error of being wrong.

The lineup protocol can be used in a variety of testing scenarios for visualizations. @GIIV suggest the choropleth map for testing spatial structure in a data set. The lineup protocol allows for flexibility in the definition of null data. In the case of spatial visualizations, it is likely that neighbors are related. To account for this, null data sets can be generated by sampling from a known model that accounts for spatial covariance.

To contrast the effectiveness of two displays, we can produce null data plots with a hidden real data plot using the different plot designs.
We can compare the time taken by participants to evaluate the same data in the different displays. We can also contrast the accuracy achieved by participants in detecting the real data plot.
-->

@@

<!--
The mean of the detection rate for choropleth map, denoted as $\mu_C$, and the hexagon tile maps, $\mu_H$ will be contrasted. This leads to the following one sided hypothesis:

$H_0$ : $\mu_H$ = $\mu_C$
$H_a$ : $\mu_H$ > $\mu_C$

The detection rate $\hat\pi$ is calculated as the amount of people that made the choice of plot that contained the real data, out of the participants who saw data plot in the lineup of the null data plots.
-->

@@

<!-- Discuss simulating the trend data -->
The real data display for each trend type was created by applying different methods to the centroid values of each geographic unit.

The North West to Soth East (NW-SE) distribution was based on the centroid longitude and latitude values. First,  the minimum latitude and longitude value of all centroids were found. Then the difference was found between the minimums, and the centroid values of all other centroids in the set of SA3 geographic units.
The following formula was applied to the difference from the minimums: $ (-2*latitude) + longitude $.

The All Cities trend model was created using the distance from the centroid of each geographic unit to the closest capital city in Australia. For each geographic unit, the distance to the closest capital city can be found when creating the hexagon tile map using the sugarbag [@sugarbag] package.
There are large geographic differences between the centroids of greater capital city areas and the rural outback areas. 
To find the geographic units closest to the capital cities, the distance to the closest capital city was rescaled to the power of 8 to help differentiate between the smallest distances.
The distances were then scaled between 0 and 1, where very small distance from the capital city were near 0, and the farthest centroid from any capital city were near to 1. The greater capital city geographic units were considered those with a scaled distance of less than 0.85, this resulted in 201 units considered "close". This was chosen to make red spots around the cities visible, even if they were not overtly noticeable in the choropleth display.

A similar process was applied to the three cities' trend model. For each of the four replicates for the Three Cities trend, a random sample of capital cities was taken from Sydney, Brisbane, Melbourne, Adelaide, Perth, and Hobart.
The same process used for the All Cities trend was applied, but only to the areas nearest to the three cities in the sample for each of the four replicates. 
For the Three cities lineups, scaling was applied separately to the values for geographic units that were and were not closest to the cities that were included in the sample. The geographic units that were close to the three cities selected were scaled to a larger maximum value so that their greatest values would appear red.

@@

One of the lineup map locations was randomly chosen to embed the real trend model data, in each of the four replicates, for the three trend models.
In this location, real data display of the trend model was added to the null values of the spatially correlated data for each lineup.

@@

For each of the 144 individual maps, the values attributed to each geographic area are rescaled to show a similar color scale from deep blue to dark red within each map.
This allowed at least one geographic unit to be colour dark blue, and at least one to be red, in every map display of every lineup.

@@

Each set of lineup data was used to produce a choropleth maps and hexagon tile maps. These matched pairs were split between Group A and Group B according to the 2 x 3 factor experimental design depicted in \ref{fig:exp-design}.

@@

### Data cleaning

<!-- A pilot study was conducted to determine whether the lineups were appropriate. All lineups in the pilot study were deemed viable as at least one participant detected the real data plot in each lineup.
The demographics of the participants showed a skew towards male participants. 
The randomness of the group allocation also resulted in more participants being allocated to Group B. Due to the allocation of lineup displays the participants all saw six Choropleth displays and six Hexagon Tile Map displays.
-->

>>>>

The paper is organised as follows. The next section discusses the background of geographic data display and visual inference procedures. The [Methodology] Section describes the methods for conducting the experiment and analysing the results. The results are summarized in Section [Results].

Background
=============

@@

@@

@@

@@

@@

Each set of lineup data was used to produce a choropleth map lineup and hexagon tile map lineup. These matched pairs were split between Group A and Group B according to the 2 x 3 factor experimental design depicted in \ref{fig:exp-design}.

<!-- Discuss simulating the trend data -->
For each trend model, the real data display was created by manipulating the centroid values of each geographic unit.

The North West to South East (NW-SE) distribution was created using a linear equation of the centroid longitude and latitude values.

The All Cities trend model was created using the distance from the centroid of each geographic unit to the closest capital city in Australia. The distance to the closest capital city for each geographic unit was found when creating the hexagon tile map using the sugarbag [@sugarbag] package.
201 Statistical Areas were considered greater capital city areas, the values of these areas were increased to create red clusters. They were chosen to make clusters around the cities visible, even if they were not overtly noticeable in the choropleth display.

A similar selection process was applied to the Three Cities' trend model. However, for each of the four replicates for the Three Cities trend, a random sample of capital cities was taken from Sydney, Brisbane, Melbourne, Adelaide, Perth, and Hobart. Only values of the areas nearest to the three cities were increased to create clusters.

@@

One of the lineup map locations was chosen to embed the real trend model data, in each of the four replicates, for the three trend models.
In this location, real data display of the trend model was added to the null values of the spatially correlated data for each lineup. This location, the associated trend values and the null data were used to create both a choropleth and hexagon tile map display.
The choropleth and hexagon lineups were split between group A and B, according to the experimental design shown in \ref{fig:exp-design}.
The location was chosen from a sub sample of the 12 possible locations. The chance of reptition using resampling was introduced to prevent participants from inducing the location by elimination, the locations 1, 7, 10 and 11 were not used.

@@

For each of the 144 individual maps, the values attributed to each geographic area were rescaled to create a similar color scale from deep blue to dark red within each map.
This allowed at least one geographic unit to be colour dark blue, and at least one to be red, in every map display of every lineup.

@@

For the population related displays, the clusters in the cities appeared more red than the rest of Australia.

### Data Cleaning

The first step in the data cleaning process involved checking that survey responses collected for each participants were only included once in the data set. 
The data cleaning process also involved filtering out participants' who did not provide any choices when considering each of the twelve lineups. These participants achieved a detection rate of 0. If participants had made various plot choices for the 12 displays they saw they were still included in the dataset.

@@

### 2020-01-14 paper/paper.tex
rearrange

The paper is organised as follows. The next section discusses the background of geographic data display and visual inference procedures. The \protect\hyperlink{methodology}{Methodology} Section describes the methods for conducting the experiment and analysing the results. The results are summarized in Section \protect\hyperlink{results}{Results}.
Background
=============

@@

The real data display for each trend type was created by applying different methods to the centroid values of each geographic unit.

The North West to Soth East (NW-SE) distribution was based on the centroid longitude and latitude values. First, the minimum latitude and longitude value of all centroids were found. Then the difference was found between the minimums, and the centroid values of all other centroids in the set of SA3 geographic units.
The following formula was applied to the difference from the minimums: \$ (-2*latitude) + longitude \$.

The All Cities trend model was created using the distance from the centroid of each geographic unit to the closest capital city in Australia. For each geographic unit, the distance to the closest capital city can be found when creating the hexagon tile map using the sugarbag {[}\protect\hyperlink{ref-sugarbag}{1}{]} package.
There are large geographic differences between the centroids of greater capital city areas and the rural outback areas.
To find the geographic units closest to the capital cities, the distance to the closest capital city was rescaled to the power of 8 to help differentiate between the smallest distances.
The distances were then scaled between 0 and 1, where very small distance from the capital city were near 0, and the farthest centroid from any capital city were near to 1. The greater capital city geographic units were considered those with a scaled distance of less than 0.85, this resulted in 201 units considered ``close''. This was chosen to make red spots around the cities visible, even if they were not overtly noticeable in the choropleth display.

A similar process was applied to the three cities' trend model. For each of the four replicates for the Three Cities trend, a random sample of capital cities was taken from Sydney, Brisbane, Melbourne, Adelaide, Perth, and Hobart.
The same process used for the All Cities trend was applied, but only to the areas nearest to the three cities in the sample for each of the four replicates.
For the Three cities lineups, scaling was applied separately to the values for geographic units that were and were not closest to the cities that were included in the sample. The geographic units that were close to the three cities selected were scaled to a larger maximum value so that their greatest values would appear red.

One of the lineup map locations was randomly chosen to embed the real trend model data, in each of the four replicates, for the three trend models.
In this location, real data display of the trend model was added to the null values of the spatially correlated data for each lineup.

For each of the 144 individual maps, the values attributed to each geographic area are rescaled to show a similar color scale from deep blue to dark red within each map.
This allowed at least one geographic unit to be colour dark blue, and at least one to be red, in every map display of every lineup.

@@

Each set of lineup data was used to produce a choropleth maps and hexagon tile maps. These matched pairs were split between Group A and Group B according to the 2 x 3 factor experimental design depicted in \ref{fig:exp-design}.

@@

The taipan {[}\protect\hyperlink{ref-taipan}{18}{]} package for R was used to create the survey web application.
This structure was altered to collect responses regarding participants demographics and their survey responses.
The survey app contained three tabs, participants were first shown the demographics and consent page.
This asked participants for their Figure-Eight contributor ID, and their consent to the responses being used for analysis. The demographics collected included participants' preferred pronoun, the highest level of education achieved, their age range and whether they had lived in Australia.

@@

Each participant used the internet to access the survey.
The data transfer from the web application to the data set took place using a secure link to the googlesheet used to store results. The application connected to the googlesheet using the googlesheets {[}\protect\hyperlink{ref-sheets}{19}{]} R package when participants opened the application, and interacted again when participants chosen to submit the survey and indicate their completion. At this time it added the participant's responses to the twelve lineup displays as twelve rows of data in the googlesheet.

@@

Participants were recruited to evaluate lineups from the Figure Eight crowdsourcing platform {[}\protect\hyperlink{ref-figeight}{20}{]}.
The lineup protocol expects that the participants are uninvolved judges with no prior knowledge of the data, to avoid inadvertently affecting results. Potential participants needed to have achieved level 2 or level 3 from prior work. All participants were at least 18 years old.

@@

The likelihood of detecting the data plot in the lineup can be modelled using a linear mixed effects model.
The R {[}\protect\hyperlink{ref-RCore}{21}{]} \texttt{glmer()} function in the {[}\protect\hyperlink{ref-lme4}{22}{]} package implements generalised linear mixed effect models. The model used includes the two main effects map type and trend model, which gives the fixed effects model to be:

Computation will be done using R {[}\protect\hyperlink{ref-RCore}{21}{]}, with the \texttt{lme4} package {[}\protect\hyperlink{ref-lme4}{22}{]}.

\hypertarget{data-cleaning}{%
\subsubsection{Data cleaning}\label{data-cleaning}}

@@

The authors would like to thank the Australian Cancer Atlas team for discussions regarding alternative spatial visualizations, and Professor Kerrie Mengersen and Dr Earl Duncan for regular meetings filled with suggestions and comments. Mitchell O'Hara-Wild was a co-developer of the taipan {[}\protect\hyperlink{ref-taipan}{18}{]} R package for image tagging, used as the base for the web app constructed to collect participant evaluations of lineups. We are thankful for the NUMBATs (Non-Uniform Monash Business Analytics Team) for participating in the pilot study that helped to assess the experimental design and determine an appropriate sample size for the study.

@@

The analysis of the work was completed in R {[}\protect\hyperlink{ref-RCore}{21}{]} with the use of the following packages:

\begin{itemize}
\tightlist
\item
  For document creation: rmarkdown {[}\protect\hyperlink{ref-rmarkdown}{23}{]}, rticles {[}\protect\hyperlink{ref-rticles}{24}{]}, knitr {[}\protect\hyperlink{ref-knitr}{25}{]}.
\item
  For lineup creation and data analysis: tidyverse {[}\protect\hyperlink{ref-tidyverse}{26}{]}, nullabor {[}\protect\hyperlink{ref-nullabor}{27}{]}, ggthemes {[}\protect\hyperlink{ref-ggthemes}{28}{]}, RColorBrewer {[}\protect\hyperlink{ref-RColorBrewer}{29}{]}.
\item
  For image displays: cowplot {[}\protect\hyperlink{ref-cowplot}{30}{]}, png {[}\protect\hyperlink{ref-png}{31}{]}, grid {[}\protect\hyperlink{ref-grid}{32}{]}.
\item
  For modelling and presentation of models: gstat {[}\protect\hyperlink{ref-gstat}{17}{]}, lme4 {[}\protect\hyperlink{ref-lme4}{22}{]}, kableExtra {[}\protect\hyperlink{ref-kableExtra}{33}{]}.
\end{itemize}

@@

\leavevmode\hypertarget{ref-taipan}{}%
18. Kobakian S, O'Hara-Wild M (2018) Taipan: Tool for annotating images in preparation for analysis.

\leavevmode\hypertarget{ref-sheets}{}%
19. Bryan J, Zhao J (2018) Googlesheets: Manage google spreadsheets from r.

\leavevmode\hypertarget{ref-figeight}{}%
20. (2019) The essential high-quality data annotation platform. Figure Eight

\leavevmode\hypertarget{ref-RCore}{}%
21. R Core Team (2019) R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria

\leavevmode\hypertarget{ref-lme4}{}%
22. Bates D, Mächler M, Bolker B, Walker S (2015) Fitting linear mixed-effects models using lme4. Journal of Statistical Software 67:1--48

>>>>

The paper is organised as follows. The next section discusses the background of geographic data display and visual inference procedures. The \protect\hyperlink{methodology}{Methodology} Section describes the methods for conducting the experiment and analysing the results. The results are summarized in Section \protect\hyperlink{results}{Results}.

\hypertarget{background}{%
\section{Background}\label{background}}

@@

Each set of lineup data was used to produce a choropleth map lineup and hexagon tile map lineup. These matched pairs were split between Group A and Group B according to the 2 x 3 factor experimental design depicted in \ref{fig:exp-design}.

For each trend model, the real data display was created by manipulating the centroid values of each geographic unit.

The North West to South East (NW-SE) distribution was created using a linear equation of the centroid longitude and latitude values.

The All Cities trend model was created using the distance from the centroid of each geographic unit to the closest capital city in Australia. The distance to the closest capital city for each geographic unit was found when creating the hexagon tile map using the sugarbag {[}\protect\hyperlink{ref-sugarbag}{1}{]} package.
201 Statistical Areas were considered greater capital city areas, the values of these areas were increased to create red clusters. They were chosen to make clusters around the cities visible, even if they were not overtly noticeable in the choropleth display.

A similar selection process was applied to the Three Cities' trend model. However, for each of the four replicates for the Three Cities trend, a random sample of capital cities was taken from Sydney, Brisbane, Melbourne, Adelaide, Perth, and Hobart. Only values of the areas nearest to the three cities were increased to create clusters.

One of the lineup map locations was chosen to embed the real trend model data, in each of the four replicates, for the three trend models.
In this location, real data display of the trend model was added to the null values of the spatially correlated data for each lineup. This location, the associated trend values and the null data were used to create both a choropleth and hexagon tile map display.
The choropleth and hexagon lineups were split between group A and B, according to the experimental design shown in \ref{fig:exp-design}.
The location was chosen from a sub sample of the 12 possible locations. The chance of reptition using resampling was introduced to prevent participants from inducing the location by elimination, the locations 1, 7, 10 and 11 were not used.

For each of the 144 individual maps, the values attributed to each geographic area were rescaled to create a similar color scale from deep blue to dark red within each map.
This allowed at least one geographic unit to be colour dark blue, and at least one to be red, in every map display of every lineup.

@@

For the population related displays, the clusters in the cities appeared more red than the rest of Australia.

\hypertarget{data-cleaning}{%
\subsubsection{Data Cleaning}\label{data-cleaning}}

The first step in the data cleaning process involved checking that survey responses collected for each participants were only included once in the data set.
The data cleaning process also involved filtering out participants' who did not provide any choices when considering each of the twelve lineups. These participants achieved a detection rate of 0. If participants had made various plot choices for the 12 displays they saw they were still included in the dataset.

The likelihood of detecting the data plot in the lineup can be modelled using a linear mixed effects model.
The R {[}\protect\hyperlink{ref-RCore}{18}{]} \texttt{glmer()} function in the {[}\protect\hyperlink{ref-lme4}{19}{]} package implements generalised linear mixed effect models. The model used includes the two main effects map type and trend model, which gives the fixed effects model to be:

Computation will be done using R {[}\protect\hyperlink{ref-RCore}{18}{]}, with the \texttt{lme4} package {[}\protect\hyperlink{ref-lme4}{19}{]}.

@@

The taipan {[}\protect\hyperlink{ref-taipan}{20}{]} package for R was used to create the survey web application.
This structure was altered to collect responses regarding participants demographics and their survey responses.
The survey app contained three tabs, participants were first shown the demographics and consent page.
This asked participants for their Figure-Eight contributor ID, and their consent to the responses being used for analysis. The demographics collected included participants' preferred pronoun, the highest level of education achieved, their age range and whether they had lived in Australia.

@@

Each participant used the internet to access the survey.
The data transfer from the web application to the data set took place using a secure link to the googlesheet used to store results. The application connected to the googlesheet using the googlesheets {[}\protect\hyperlink{ref-sheets}{21}{]} R package when participants opened the application, and interacted again when participants chosen to submit the survey and indicate their completion. At this time it added the participant's responses to the twelve lineup displays as twelve rows of data in the googlesheet.

@@

Participants were recruited to evaluate lineups from the Figure Eight crowdsourcing platform {[}\protect\hyperlink{ref-figeight}{22}{]}.
The lineup protocol expects that the participants are uninvolved judges with no prior knowledge of the data, to avoid inadvertently affecting results. Potential participants needed to have achieved level 2 or level 3 from prior work. All participants were at least 18 years old.

@@

@@

The authors would like to thank the Australian Cancer Atlas team for discussions regarding alternative spatial visualizations, and Professor Kerrie Mengersen and Dr Earl Duncan for regular meetings filled with suggestions and comments. Mitchell O'Hara-Wild was a co-developer of the taipan {[}\protect\hyperlink{ref-taipan}{20}{]} R package for image tagging, used as the base for the web app constructed to collect participant evaluations of lineups. We are thankful for the NUMBATs (Non-Uniform Monash Business Analytics Team) for participating in the pilot study that helped to assess the experimental design and determine an appropriate sample size for the study.

@@

The analysis of the work was completed in R {[}\protect\hyperlink{ref-RCore}{18}{]} with the use of the following packages:

\begin{itemize}
\tightlist
\item
  For document creation: rmarkdown {[}\protect\hyperlink{ref-rmarkdown}{23}{]}, rticles {[}\protect\hyperlink{ref-rticles}{24}{]}, knitr {[}\protect\hyperlink{ref-knitr}{25}{]}.
\item
  For lineup creation and data analysis: tidyverse {[}\protect\hyperlink{ref-tidyverse}{26}{]}, nullabor {[}\protect\hyperlink{ref-nullabor}{27}{]}, ggthemes {[}\protect\hyperlink{ref-ggthemes}{28}{]}, RColorBrewer {[}\protect\hyperlink{ref-RColorBrewer}{29}{]}.
\item
  For image displays: cowplot {[}\protect\hyperlink{ref-cowplot}{30}{]}, png {[}\protect\hyperlink{ref-png}{31}{]}, grid {[}\protect\hyperlink{ref-grid}{32}{]}.
\item
  For modelling and presentation of models: gstat {[}\protect\hyperlink{ref-gstat}{17}{]}, lme4 {[}\protect\hyperlink{ref-lme4}{19}{]}, kableExtra {[}\protect\hyperlink{ref-kableExtra}{33}{]}.
\end{itemize}

@@

\leavevmode\hypertarget{ref-RCore}{}%
18. R Core Team (2019) R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria

\leavevmode\hypertarget{ref-lme4}{}%
19. Bates D, Mächler M, Bolker B, Walker S (2015) Fitting linear mixed-effects models using lme4. Journal of Statistical Software 67:1--48

\leavevmode\hypertarget{ref-taipan}{}%
20. Kobakian S, O'Hara-Wild M (2018) Taipan: Tool for annotating images in preparation for analysis.

\leavevmode\hypertarget{ref-sheets}{}%
21. Bryan J, Zhao J (2018) Googlesheets: Manage google spreadsheets from r.

\leavevmode\hypertarget{ref-figeight}{}%
22. (2019) The essential high-quality data annotation platform. Figure Eight
